### 2016-07-18 content.tex

\title{%
  Verifying Demonstrations
}
\author[D.~Bosk et al.]{%
  Daniel Bosk\inst{1}
  \and
  Sébastien Gambs\inst{2}
  \and
  Sonja Buchegger\inst{1}
}
\institute[KTH, UQAM]{%
  School of Computer Science and Communication\\
  KTH Royal Institute of Technology, Stockholm\\
  \texttt{\{dbosk,buc\}@kth.se}
  \and
  Université de Quebec à Montreal\\
  \texttt{sgambs@uqam.ca}
}

\maketitle

\mode*

\begin{abstract}
  \dots

  \keywords{\dots}
\end{abstract}

\input{intro.tex}
\input{AdaptingPROPS.tex}
\input{SystemModel.tex}
\input{primitives.tex}

\subsubsection*{Acknowledgements}

This work was funded by the Swedish Foundation for Strategic Research grant SSF 
FFL09-0086 and the Swedish Research Council grant VR 2009-3793.

\printbibliography{}