### 2017-06-01 DataAuthenticity.tex
Adds material from Protesting Chapter
Adapts material so that it compiles, but no editing.

\section{Data authenticity}
\label{DataAuthenticity}

The problem of authentically associating data with a physical event is 
difficult.
We can essentially divide it into the following requirements:
\begin{requirements}[A]
  \item\label{CreatedAfterStart} Prove that the data was created after the 
    start of the event.
  \item\label{CreatedBeforeEnd} Prove that the data was created before the end 
    of the event.
  \item\label{SpatiallyRelated} Prove that the data is spatially related to the 
    physical location of the event.
\end{requirements} %TODO: A4 accurate/not fabricated?
\Cref{CreatedBeforeEnd,CreatedAfterStart} together bind the data to the time of 
the event whereas \cref{SpatiallyRelated} binds the data spatially to the 
event.

Now, consider the scenario of Alice taking a photo during the demonstration and 
posting it online.
What can we say about this photo?
First, we can say that it was created before we viewed it, so 
\cref{CreatedBeforeEnd} is fulfilled if we view it in relation to the event.
If it was submitted to a service that we trust, e.g.\ a blockchain, then we can 
also trust the time-stamp of the service for fulfilling 
\cref{CreatedBeforeEnd}.
Furthermore, we can also consider it spatially related to the physical location
(\cref{SpatiallyRelated}) if we can convince ourselves that the photo is 
depicting the physical location and not any kind of \enquote{reconstruction}, 
e.g.\ it is not computer generated or a photo of a similarly looking location.

\Cref{CreatedAfterStart} is more difficult to achieve.
In the above scenario, there is nothing that prevents Alice from submitting an 
older photo which is spatially related to the event.
In the security field \cref{CreatedAfterStart} is captured by a property called 
\emph{freshness}.
The freshness property is usually achieved by requiring that the data depend on
an unpredictable value.
The unpredictable value is commonly a value chosen randomly by the verifier, 
but there are ways to do this without the verifier too --- the basic 
requirement is that it is not under the prover's control, i.e.\ Alice in this 
scenario.
The idea is illustrated by the following example: we require Alice to include 
the front-page of a particular newspaper in the photo, since the exact 
front-page is difficult for Alice to predict in advance.
This particular example would of course not work in practice, since Alice can 
manipulate the photo afterwards to superimpose the required newspaper 
front-page\footnote{%
  There is, however, a subfield of digital forensics that works with image 
  manipulation detection, so there are methods that would prevent at least 
  Alice's easiest manipulation attempts.
}.
But this illustrates another important requirement for freshness: that the 
operation binding the unpredicatble value to the data must be difficult to redo 
later to change the value.