### 2017-07-13 ProtestVerif-slides.tex
Fixes affil and abstract

\documentclass[ignorenonframetext]{beamer}
%\documentclass[handout,ignorenonframetext]{beamer}
%\usepackage{pgfpages}
%\pgfpagesuselayout{4 on 1}[a4paper,border shrink=5mm]

\usetheme{Berlin}
\setbeamercovered{transparent}
\setbeamertemplate{bibliography item}[text]

\setbeamertemplate{footline}%{miniframes theme}
{%
  \begin{beamercolorbox}[colsep=1.5pt]{upper separation line foot}
  \end{beamercolorbox}
  \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
    leftskip=.3cm,rightskip=.3cm plus1fil]{author in head/foot}%
    \leavevmode{\usebeamerfont{author in head/foot}\insertshortauthor}%
    \hfill%
    {\usebeamerfont{institute in head/foot}\usebeamercolor[fg]{institute in head/foot}\insertshortinstitute}%
  \end{beamercolorbox}%
  \begin{beamercolorbox}[ht=2.5ex,dp=1.125ex,%
    leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    {\usebeamerfont{title in head/foot}\insertshorttitle} \hfill     \insertframenumber%
  \end{beamercolorbox}%
  \begin{beamercolorbox}[colsep=1.5pt]{lower separation line foot}
  \end{beamercolorbox}
}

\AtBeginSection[]{%
  \begin{frame}<beamer>
    \tableofcontents[currentsection]
  \end{frame}
}

\let\email\texttt

\input{preamble.tex}
\input{preamble-slides.tex}

\author[D.~Bosk et al.]{%
  Daniel Bosk\inst{1}
  \and
  Sonja Buchegger\inst{1}
  \and
  Sébastien Gambs\inst{2}
}
\institute[KTH, UQAM]{%
  \inst{1} School of Computer Science and Communication\\
  KTH Royal Institute of Technology, Stockholm\\
  \texttt{\{dbosk,buc\}@kth.se}
  \and
  \inst{2} Université de Quebec à Montreal\\
  \texttt{sgambs@uqam.ca}
}

\begin{document}
\mode<all>
\input{contents.tex}
\mode*
\end{document}