### 2017-07-13 abstract.tex

Alice organizes a protest for some change in government.
The protest is held in some location(s).
Alice wants to estimate the number of participants to prove certain support for
her change.
Current methods have wide margins for error.
In addition, Alice's incentive is to estimate the number of participants as 
high as possible whereas the government's incentive is to estimate the number 
of participants as low as possible.

We are interested in combining location proofs with electronic voting.
Since the parties' have incentives to cheat, we require transparency: we desire 
a decentralized scheme that provides the verifiability and privacy properties 
of electronic voting.
The scheme must be scalable to handle large protests.

\keywords{%
  crowd-counting; protesting; electronic voting; location proofs; privacy
}