### 2017-07-20 RelatedWork.tex
Adds WIP on related work

\mode*
\section{Related work}

\Textcite{2016DemonstrationsInSeoul} discusses various approaches to crowd 
counting.

\Textcite{HowToEstimateCrowdSize} \dots

\Textcite{WifiCrowdCounting} \dots

\Textcite{CVCrowdCounting} \dots

\Textcite{NNCrowdCounting} \dots

The most closely related work is \textcite{CrowdCount.org}.
This is a web service which lets Alice create an event and Bob can submit his 
location to register that he is in Alice's event.
There is no verification, i.e.\ the service must be trusted, nor anything 
preventing Bob from submitting twice.
The service also requires an Internet connection during the event to register 
as a participant.
This can make it difficult to use if e.g.\ Alice has organized a protest 
against government, who shuts down the cellular network and Internet backbone 
as a means to censor the protest.