### 2017-07-25 conclusions.tex

\mode*
\section{Conclusions}

In general, none of the related works fulfils the authenticity requirements 
(\cref{%
  CreatedAfterStart,CreatedBeforeEnd,SpatiallyRelated%
}).
Some of the schemes provide some limited verifiability, i.e.\ mostly universal 
verifiability (\cref{UniversalVerif}), but not individual verifiability 
(\cref{IndividualVerif}) nor eligibility verifiability 
(\cref{EligibilityVerif}).

\begin{frame}<presentation>
  \begin{remark}
    \begin{itemize}
      \item All current methods lack authenticity and verifiability.
      \item Some provide universal verifiability (V\ref{UniversalVerif}).
    \end{itemize}
  \end{remark}
\end{frame}