### 2017-07-28 primitives.tex
Changes solution section to primitives

\mode*

\section{Useful Primitives}
\label{UsefulPrimitives}

\dots

From a computational-complexity perspective, operations that must be done on 
the phone must be cheap, other operations may be expensive.