### 2017-07-28 primitives.tex
Changes section title for primitives

\mode*

\section{Primitives}
\label{Primitives}

\dots

From a computational-complexity perspective, operations that must be done on 
the phone must be cheap, other operations may be expensive.