### 2017-08-11 conclusions.tex

\mode*
\section{Conclusions}

In general, none of the related works fulfils the eligibility verifiability 
\cref{%
  CountOnce,%
  CreatedAfterStart,CreatedBeforeEnd,SpatiallyRelated,%
}.
To some extent, some techniques do fulfil \cref{DesignatedEvent}, since people 
usually use placards in the protests.
Also, some schemes provide universal verifiability (\cref{UniversalVerif}), but 
not individual verifiability (\cref{IndividualVerif}).
If they would provide individual verifiability, they would likely violate the 
privacy requirements.
And finally, most of the current methods are estimates, with quite some error 
margins, so they cannot give the correct participation count.

\begin{frame}<presentation>
  \begin{remark}
    \begin{itemize}
      \item All current methods lack most of verifiability.
      \item Some provide universal verifiability (V\ref{UniversalVerif}).

        \pause

      \item Most cannot estimate the correct participation count.
    \end{itemize}
  \end{remark}
\end{frame}