### 2017-08-23 security.tex
Adds stealing participation to security section

\mode*

\section{Security}

\subsection{Stealing participation}

The Eve (possibly with Rusuk's aid) wants to arrange a counter-protest in Eve's 
favour.
One way to get a higher participation count is to use the same identifier as 
Alice's protest.
Since the identifier for the protest is the hash value of the manifest, Eve 
must find a second pre-image for the hash function.

\mode<presentation>{%
\begin{frame}
  \begin{itemize}
    \item Eve wants to arrange a counter-protest.
    \item She wants to steal some of Alice's participants' proofs.
    \item To do this she must have an identifier that matches.
    \item I.e.\ she must find a second pre-image for the hash function.
  \end{itemize}
\end{frame}
}

% XXX Rephrase identifier problem as a game
\begin{proposition}
  Given \(x\) and \(H(x) = y\), the probability of finding an \(x'\) such that 
  \(H(x') = y\) is \(\frac{1}{\lambda-1}\).
\end{proposition}