### 2017-08-27 fig/tposet.tex
Fixes arrows in poset graph
The arrows should be the other direction, since it links backwards, not
forwards.

\begin{tikzpicture}[->,auto,node distance=1.5cm,thick,
  proof node/.style={rectangle,draw},
  time node/.style={circle,draw}]

  \node (prev) {\dots};
  \node[time node] (ts) [right of=prev] {$t_s$};
  \node[proof node] (pA) [right of=ts] {$p_A$};
  \node[proof node] (pB) [below of=pA] {$p_B$};
  \node[time node] (te) [right of=pA] {$t_e$};
  \node (succ) [right of=te] {\dots};

  \path
  (succ) edge node {} (te)
  (te) edge node {} (pB)
  (pB) edge node {} (ts)
  (te) edge node {} (pA)
  (pA) edge node {} (ts)
  (ts) edge node {} (prev);

\end{tikzpicture}