### 2017-08-29 conclusions.tex
Adds clarifications, improves slides

\mode*
\section{Conclusions}

\subsection<presentation>{Related work}

In general, none of the related works fulfils the eligibility verifiability 
\cref{%
  CountOnce,%
  CreatedAfterStart,CreatedBeforeEnd,SpatiallyRelated,%
}.
To some extent, some techniques do fulfil \cref{DesignatedEvent}, since people 
usually use placards in the protests.
Also, some schemes provide universal verifiability (\cref{UniversalVerif}), but 
not individual verifiability (\cref{IndividualVerif}).
If they would provide individual verifiability, they would likely violate the 
privacy requirements.
And finally, most of the current methods are estimates, with quite some error 
margins, so they cannot give the correct participation count.

\begin{frame}<presentation>
  \begin{block}{Verifiability}
    \begin{itemize}
      \item All current methods lack most of verifiability.
      \item Some can provide universal verifiability.
    \end{itemize}
  \end{block}

  \pause{}

  \begin{block}{Correctness}
    \begin{itemize}
      \item Most cannot estimate the correct participation count.
      \item Some can (theoretically) give an upper bound.
    \end{itemize}
  \end{block}
\end{frame}

\subsection{Open problems}

\begin{frame}
  \frametitle{Sybil}
  \begin{question}
    How to solve Sybil, can we piggyback on existing \acp{PKI}?
    E.g.\ through~\cite{Cinderella}.
  \end{question}

  \begin{question}
    Can we otherwise bind identifiable keys to unique ring signature keys?
  \end{question}

  \begin{question}
    Can we achieve receipt freeness (deniability) and protect from Sybil?
  \end{question}
\end{frame}

\begin{frame}
  \frametitle{Anonymous communication}
  \begin{question}
    How do we submit the proofs to storage?
  \end{question}

  \mode<presentation>{%
    \begin{remark}
      Without anonymous submission we don't get proof privacy, thus no 
      receipt freeness either.
    \end{remark}
  }
\end{frame}

\begin{frame}
  \frametitle{Blockchains, distributed systems}
  \begin{question}
    \begin{itemize}
      \item Malicious nodes want to provide Alice with a view that her proof is 
        included.
      \item Everyone else is provided a view without Alice.
      \item How can we ensure a consistent view for everyone?
    \end{itemize}
  \end{question}

  \begin{question}
    Can we do blockchain branch and merge?
  \end{question}
\end{frame}