### 2017-09-20 primitives.tex
Changes section title of security analysis

\section{Building blocks}%
\label{Primitives}\label{BuildingBlocks}

\dots

\subsection{Zero-knowledge proofs}

Briefly describe zero-knowledge proofs and what type of properties can be proved 
efficiently using them, e.g.\ algebraic structures but not one-way functions.

\dots

\subsection{Location proofs}

Briefly describe location proofs and distance bounding protocols.
Describe the different desirable properties for distance bounding:
\begin{itemize}
  \item Mafia-fraud resistance,
  \item terrorist-fraud resistance,
  \item impersonation-fraud resistance,
  \item distance-fraud resistance.
\end{itemize}