### 2018-02-22 paper/BuildingBlocks.tex
Rearranges Building Blocks section

\section{Building blocks}%
\label{Primitives}\label{BuildingBlocks}

\dots

\input{anon-cred.tex}
\input{DistanceBounding.tex}
\input{LocationProofs.tex}
\input{TimeStamp.tex}