### 2018-02-26 paper/MANETs.tex

%@sbuc: Can you write a few words about MANETs (file MANETs.tex)?
%More specifically, why we can assume there is a way for the
%protesters to communicate locally at the protest without relying on the
%Internet. Also mention potential privacy problems at this layer, e.g.\
%MAC-address tracking etc. which is outside of our protocol (this should
%cover part of issue #58).

Since the cellular network can be shut down to keep protesters from
accessing the Internet, making phone calls or texting, we require a
different means of communications between protesters. This can be
accomplished by Bluetooth or WiFi, as demonstrated by
Briar~\cite{Briar} and FireChat~\cite{FireChat}, two examples of apps
for communication during protests via wireless mesh networking. An
alternative, although originally designed to work within 5G cellular
networks, is device-to-device communication (D2D)~\cite{D2D}. Specifically, the
outband and autonomous version D2D fits our scenario thanks to using
unlicensed spectrum and working without cellular coverage.