### 2018-02-27 paper/TimeStamp.tex
Adds description of blockchain, TS

\subsection{Time-stamping service, or, blockchains}%
\label{StorageProperties}

\NewCryptoScheme{\TS}{TS}
\NewAlgorithm{\TSget}{\TS.\!Get}
\NewAlgorithm{\TSstamp}{\TS.\!Stamp}
\NewAlgorithm{\TSverify}{\TS.\!Verify}
\NewAlgorithm{\TStime}{\TS.\!Time}

We need a time-stamping service, \(\TS\), with algorithms \(\TSget\), 
\(\TSstamp\) and \(\TStime\) such that
\begin{itemize}
  \item \(\rho \gets \TSget\) yields a value \(\rho\) at time \(t\), \(\rho\) is 
    difficult to guess before time \(t\) and \(\TStime[\rho] = t\);
  \item \(\pi\gets \TSstamp[x]\) yields a value \(\pi\) at time \(t\) such that 
    \(\TSverify[x, \pi]\to 1\) and \(\TStime[\pi] = t\).
\end{itemize}

With these building blocks we can ensure that a message \(m\) was created within 
the time interval \(\interval{t_0}{t_1}\).
After time \(t_0\), request \(\rho_{t_0}\gets \TSget\).
Before time \(t_1\), submit \(h\gets \Hash[m, \rho_{t_0}]\) to the time-stamping 
service to get \(\pi_{t_1}\gets \TSstamp[h]\).

Now we can use \((\rho_{t_0}, m, \pi_{t_1})\) to prove that \(m\) was created 
within the time interval \(\interval{t_0}{t_1}\).
The verifier computes \(h'\gets \Hash[m, \rho_{t_0}]\) and checks whether 
\(\TSverify[h', \pi_{t_1}] = 1\) and \(\TStime[\rho_{t_0}] = t_0\land 
  \TStime[\pi_{t_1}] = t_1\).

We can instantiate \(\TS\) using a blockchain.
The \(\TSstamp[x]\) algorithm will simply include \(x\)

\(\TSget\) will return the hash of the current head of the chain.
This value is difficult to predict since it depends on all submitted 
transactions.
This forces us to make the assumption that the adversary cannot control 
\emph{all} transactions that are submitted to the blockchain.
As long as one non-adversarial transaction is submitted it will be 
unpredictable.

We will use the storage to provide the temporal eligibility property.
\Cref{CreatedBeforeEnd,CreatedAfterStart} requires a \emph{partially ordered 
  set}\footnote{%
  A relation \(\preceq\) which is reflexive, antisymmetric and transitive.
} of objects.
If some objects in the set relate to known points in time, then the partial 
order relates the data to the time of the event.
This allows us to \emph{verify the data temporally}.
One primitive that fulfils these requirements is a blockchain.
There are also other structures, e.g.\ a directed graph that 
converges~\cite{BlockchainFreeCryptocurrencies}, that also provides the required 
properties.
However, we will use the blockchain in our discussion for simplicity.

There are five properties that we need from the storage system, all of which 
makes a blockchain-like structure suitable.
First, we need immutability.
I.e.\ once we commit something to the blockchain it will remain there.
We will use this to ensure verifiability, the data must remain to be verifiable.

Second, we need a general form of time-stamping.
I.e.\ we must be able to relate commitments to time, e.g.\ in 
Bitcoin~\cite{Bitcoin} a new block appears approximately every 10 minutes.
This coarse-grained form of time-stamping is enough for our purposes.
We will use this property to ensure that proofs are created before a certain 
point in time.

Third, we need an unpredictable number whose time of publication can be 
verified.
For this property, we will use the head of the blockchain.
The hash value of the head of the blockchain at a given time will be difficult 
to predict ahead of time.
%(See \cref{SecurityAnalysis} for a security analysis.)
We want to use this to ensure that proofs are created after a certain point in 
time.

Fourth, once we have committed some data, we do not need to keep the data 
itself, but we can store some confirmation value instead.
E.g.\ using blockchains, we can store the hash of the block where our data was 
committed, while this block remains we are sure that our data remains committed 
--- even if we no longer remember what our data looked like.
We will use this property to approximate the property of receipt freeness.

Finally, if the storage is decentralized we do not have to trust any central 
authority, which is the usual case for blockchains.