### 2018-02-28 paper/TimeStamp.tex
Fixes the blockchain section

\subsection{Time-stamping service, or, blockchains}%
\label{StorageProperties}

\NewCryptoScheme{\TS}{TS}
\NewAlgorithm{\TSget}{\TS.\!Get}
\NewAlgorithm{\TSstamp}{\TS.\!Stamp}
\NewAlgorithm{\TSverify}{\TS.\!Verify}
\NewAlgorithm{\TStime}{\TS.\!Time}

We need a time-stamping service, \(\TS\), with algorithms \(\TSget\), 
\(\TSstamp\) and \(\TStime\) such that
\begin{itemize}
  \item \(\rho \gets \TSget\) yields a value \(\rho\) at time \(t\), \(\rho\) is 
    difficult to guess before time \(t\) and \(\TStime[\rho] = t\);
  \item \(\pi\gets \TSstamp[x]\) yields a value \(\pi\) at time \(t\) such that 
    \(\TSverify[x, \pi]\to 1\) and \(\TStime[\pi] = t\).
\end{itemize}

With these building blocks we can ensure that a message \(m\) was created within 
the time interval \(\interval{t_0}{t_1}\).
After time \(t_0\), request \(\rho_{t_0}\gets \TSget\).
Before time \(t_1\), submit \(h\gets \Hash[m, \rho_{t_0}]\) to the time-stamping 
service to get \(\pi_{t_1}\gets \TSstamp[h]\).

Now we can use \((\rho_{t_0}, m, \pi_{t_1})\) to prove that \(m\) was created 
within the time interval \(\interval{t_0}{t_1}\).
The verifier computes \(h'\gets \Hash[m, \rho_{t_0}]\) and checks whether 
\(\TSverify[h', \pi_{t_1}] = 1\) and \(\TStime[\rho_{t_0}] = t_0\land 
  \TStime[\pi_{t_1}] = t_1\).

We can instantiate \(\TS\) using a blockchain.
The \(\TSstamp[x]\) algorithm will simply include \(x\) in the blockchain.
\(\TSget\) will return the hash of the current head of the chain.
This value is difficult to predict since it depends on all submitted 
transactions.
This forces us to make the assumption that the adversary cannot control 
\emph{all} transactions that are submitted to the blockchain.
As long as one non-adversarial transaction is submitted it will be 
unpredictable.
The blockchain must also be continuously extended, \eg as in
Bitcoin~\cite{Bitcoin}, where a new block appears approximately every 10 
minutes.
This coarse-grained form of time-stamping is enough for our purposes.

There are some additional properties that we need from \(\TS\).
First, we need immutability, \ie once we commit something (through \(\TSstamp\)) 
it will remain there.
This is usually true for blockchains.
We will use this to ensure verifiability, the data must remain to be verifiable.

Second, once we have committed some data, we do not need to keep the data 
itself, but we can store some confirmation value instead.
\Eg using blockchains, we can store the hash of the block where our data was 
committed, while this block remains we are sure that our data remains committed 
--- even if we no longer remember what our data looked like.
We will use this property to approximate the property of receipt freeness.