### 2018-03-01 paper/BuildingBlocks.tex
coincoin

\section{Building blocks}%
\label{Primitives}\label{BuildingBlocks}

In this section, we will briefly review the primitives that form the building blocks of \PRIVO.

%Seb: I am not sure that to explain hash functions as they are really basic primitives
\subsection{Collision-resistant hash functions}

\NewAlgorithm{\Hash}{H}
\NewAlgorithm{\algA}{A}

One primitive that we need is a collision-resistant hash function, \Hash.

\begin{definition}[Collision resistance]
  Let \(\lambda\) denote the security parameter.
  We say that a hash function, \(\Hash\), is collision resistant if for any \ac{PPT} algorithm \(\algA\) we have that
  \begin{multline*}
    \Pr\left[ x_1, x_2\gets \algA[1^\lambda] :
      \right. \\ \left.
      \Hash[x_1] = \Hash[x_2] \land x_1\neq x_2 \right] < \negl[\lambda].
  \end{multline*}
\end{definition}

\(\Hash\) can be instantiated with any algorithm from the SHA-2 or SHA-3 families.

\input{ZKPK.tex}

\input{anon-cred.tex}

%\input{LocationProofs.tex}
%Seb: I have mode the location proofs to the

\input{DistanceBounding.tex}

\input{DB-anon-cred.tex}

\input{TimeStamp.tex}