### 2018-05-23 paper/LocationProofs.tex
Edits related work

\subsection{Location proofs}

Some \ac{LBS} only grant access to resources to users located at a particular 
location,  thus raising the issue of verifying the position claimed by a 
particular user. In most current schemes, the location of a user/device is 
determined by the device itself (\eg through GPS) and forwarded to the \ac{LBS} 
provider. One of the main drawback of this approach is that a user can cheat by 
having his device transmitting a false location. Therefore, it is possible for a 
user to be inappropriately granted access to a particular resource while being 
thousands of kilometers away.

To counter this threat, \iac{LBS} should ask the requesting device to formally 
prove that it really is at the claimed location. This notion can be formalized 
through the concept of \emph{location proof}. In a nutshell, a location proof is 
a digital certificate attesting that someone was at a particular location at a 
specific moment in time. A location proof architecture is a system by which 
users can obtain location proofs from neighboring witnesses (\eg trusted access 
points or other users) that can later be shown to verifiers who can check the 
validity of a particular proof. Most of the existing approaches to location 
proofs require the prover and the witnesses to disclose their identities, thus 
raising many privacy issues such as the possibility of tracing the movements of 
users of the location proof architecture.
\seb{to do add the corresponding reference}
However, some location proofs systems, such as \ac{PROPS}, exists that provides 
strong privacy guarantees along with the possibility of verifying the claim of 
the location. \PRIVO share some similarities with \ac{PROPS}, although their 
objective is quite different as we aim at verifying a global property of the 
population (\ie crowd estimation) in contrast to checking the location claim 
made by a user, which is an individual property.