### 2018-05-25 paper/DistanceBounding.tex

\subsection{Distance bounding}%
\label{DistanceBounding}

\Ac{DB} protocols were first suggested by \citet{DistanceBounding} to prevent 
relay attacks in contactless communications in which the adversary forwards a 
communication between a prover and a possibly far-away verifier to authenticate. 
These attacks cannot be prevented by cryptographic tools as they are independent 
of the semantics of the messages exchanged, so mechanisms ensuring the physical 
proximity between a verifier and a prover must be used.
\Ac{DB} protocols precisely enable the verifier to estimate an upper bound on his distance to the prover by measuring the time-of-flight of short challenge-response messages (or rounds) exchanged during time-critical phases. 
Time critical phases are complemented by slow phases during which the time is not taking into account.
At the end of a \Ac{DB} protocol, the verifier should be able to determine if the prover is legitimate and in his vicinity.

The main security requirements of \ac{DB} protocols can be summarized as follows:
\begin{itemize}
  \item \acf{DBMF}: the adversary illegitimately authenticates, possibly using 
    a far-away honest prover.
  \item\acf{DBTF}: a legitimate but malicious prover helps a 
    close-to-the-verifier accomplice to authenticate.
  \item \acf{DBDF}: a legitimate but malicious prover wants to fool the 
    verifier on the distance between them.
  \item \acf{DBDH}: similar to \ac{DBDF}, sometimes using the presence of an 
    honest prover close to the verifier.
  \item \acf{DBIF}: the adversary plays against a simplified version of the 
    protocol without any distance estimation.
\end{itemize}
There are two lines of attempts at formalizing the above properties: one by 
\citet{DB-BMV} and another by \citet{DB-DFKO}.

The majority of the existing \ac{DB} protocols are symmetric and thus requires an honest verifier.
Indeed, in this context it does not make sense to protect against the verifier 
as he can easily impersonate the prover as he has a knowledge of his secret key.
There has been less work done in the domain of asymmetric (or public-key) \ac{DB} protocols.
Unfortunately, our setting requires a public-key \ac{DB} protocol with a \emph{malicious verifier} who will try to \emph{impersonate the prover}.
The verifier might also try to track the provers and map their identities to their actions, thus we also require privacy.
This leads to the requirement of a \ac{DB} \ac{ZKPK}, or simply \ac{DBPK}, with resistance to all the frauds mentioned above as well as against this verifier 
who will attempt to impersonate the prover.
More specifically, we must combine such \iac{DB} with anonymous credentials 
(which we do in \cref{DB-anon-cred}).
The anonymous credential system will affect what properties we require of the \ac{DBPK}, (\eg that it must be \iac{PKE} for discrete logarithms).