### 2018-11-28 paper/verifiability-analysis.tex
Edits verifiability analysis, aligns it with adversary

\subsection{Individual and universal verifiability}%
\label{analysis-individual}%
\label{analysis-universal}

\Cref{IndividualVerif} requires that Alice and Bob, as participants, can verify 
that their participation proofs (\ie proof shares) are indeed included in the 
computed count.
All proof shares (\ie \(\prf_{\pid, \prtst} = (\cid, \pid, \wid, t_s, t_s', 
  t_e, t_e', l, \pi_\pid, \pi_{\wid})\)) are committed to the ledger and 
available from a public and permanent storage (see 
\cref{fig:ProtocolOverview}).
Thus, Alice and Bob can simply check that all of their proof shares are indeed 
there and the security of individual verifiability depends on the properties of 
the ledger and storage (\(\TS\), \cref{timestamp}).
For both \cref{in-system-adversary,deanonymizing-storage-adversary} we assumed 
an honest-but-curious adversary controlling \(\TS\)\footnote{%
  We note that, in general, distributed (decentralized) ledgers cannot 
  withstand a malicious \ac{ISP}.
  Such an adversary can partition the network and provide Alice and Bob with 
  different views of the ledger, thus breaking individual verifiability.
  However, this requires that the adversary \emph{can observe} Alice's and 
  Bob's channels to the ledger --- which is still not true even for 
  \cref{deanonymizing-storage-adversary}.
}.
This means that Alice can check that her proof share is indeed there.

\Cref{UniversalVerif} implies that anyone can check the result and that all 
participation proofs counted are legitimate.
As the proof shares are committed and stored publicly, anyone can
download them, verify eligibility (\ie verify \(\pi_\pid,
\pi_{\wid}\)) of the proofs and count them.
As for individual verifiability, the security of universal verifiability is 
reduced to the properties of the ledger and storage; but universal 
verifiability also depends on the eligibility verification, which we will 
discuss next.

\subsection{Eligibility verifiability}%
\label{analysis-eligibility}

\Cref{EligibilityVerif} states that anyone must be able to determine the authenticity of the relevant attributes of the data.
In \CROCUS, we have several attributes that must be verifiable: the time of 
creation (\ie, temporal eligibility, \cref{TemporallyRelated}), the physical 
location of \(\sk_P\) at creation (\ie, spatial eligibility, 
\cref{SpatiallyRelated}), recognition of two proofs originating from the same 
person (\ie, one-proof-per-person eligibility, \cref{CountOnce}) and that the 
proof is indeed designated for the event (\ie, designated-event eligibility, 
\cref{DesignatedEvent}).
It follows from \cref{analysis-individual} that the adversary cannot drop 
submitted proof shares and thus cannot decrease the count.
As indicated in \cref{in-system-adversary,deanonymizing-witness-adversary}, the 
adversary can submit proof shares as everyone else, so the adversary's only 
option is to increase the count.
We will thus let Alice pose as the adversary in this section, as she naturally 
has an incentive to increase the count.

\paragraph{Temporal eligibility}%
\label{analysis-temporal}

\Cref{TemporallyRelated} ensures freshness, as Alice cannot simply resubmit an 
old proof as a new one or create a proof in advance.

In general, to prevent replays, Alice must respond to an unpredictable 
challenge.
The challenge here is \enquote{which is the hash value of the head block of the 
  ledger at the time of the proof's creation?}.
The response is simply the hash itself (obtained through \(\TSget\)) and is 
included as \(t_s\) and \(t_s'\) in the proof share (see \cref{fig:ProofFig}).

As it is the case in the construction of ledgers, the hash value of the head 
block depends on the previous blocks.
Thus, the predictability of \(t_s\) depends on the predictability of the 
blocks, which depends on the predictability of the transactions.
Assume that Alice can control all transactions going into a block.
This means that she must find a pre-image of the form of a valid block that is 
accepted by all nodes in the distributed ledger.
\Ie it depends on the pre-image resistance of the hash function used.
Thus, with a collision resistant hash function (\(\Hash\)), Alice can predict 
\(t_s\) only with negligible probability; and it follows that, with high 
probability, \(t_s\) was published on the ledger before Alice included it in 
her proof share.

The correctness of the response must be verifiable by any verifier (verifiable 
through \(\TStime\)).
This can be done if the ledger is continuously extended at a fixed rate, \eg 
one new block every 10 minutes, as was the case for Bitcoin.

According to \cref{TemporallyRelated}, we must also prove that a proof share has not been created after a certain time.
Otherwise, Grace could argue that the proof share was created after the protest, thus defeating the purpose of our protocol.
The hash values of the proof shares are committed to the ledger, which means 
that there is a negligible probability that they were created after that.
Alice is in the same situation as above:
she would have to choose a value \(y\) in the range of the hash function 
\(\Hash\) and then find a pre-image \(x\) such that \(y = \Hash[x]\) and \(x\) 
is a valid proof for the desired protest, at the desired time.
If \(\Hash\) is collision resistant, finding \emph{any} pre-image is hard.

\paragraph{Linkability}%
\label{analysis-linkability}

\Cref{CountOnce} is required to prevent Sybil attacks, in the sense
that Alice cannot provide two (or more) participation proofs for a
specific protest and thus be counted more than once.
To do this she must create more than one pseudonym, \(\pid\).
Indeed to be counted twice, Alice must produce a \(\pid'\neq \pid\).
Due to the deterministic property of \(\ACprf\), Alice must produce a new key 
\(\sk_P'\) such that the verifier\footnote{%
  Here the verifier is either the witness during the distance bounding or the 
  verifier who tries to verify the count.
} accepts the proof
\begin{multline*}
\PK\mleft\{ (\sk_P') : \pid' = \ACprf[_{\sk_P'}][\cid] \quad \land \mright. \\
    \sigma_P'' = \mleft. \ACblind[\ACsign[_{\ssk}][\sk_P']] \mright\}
\end{multline*}
while she does not know a valid signature on \(\sk_P'\).
As a consequence, this is reduced to the security of the \(\AC\) scheme and how 
often Alice can get a valid signature on a secret key from the \ac{CA} --- 
either by requesting a new key, \eg request a new passport or ID-card, or 
illicitly use someone else's.

\paragraph{Spatial eligibility}%
\label{analysis-spatial}

\Cref{SpatiallyRelated} is achieved by having a witness vouch anonymously that Alice was indeed on the location when the proof share was created.
To realize this, we must ensure that Alice cannot forge witness signatures.

In the trust-free case (see \cref{ProtocolVerification}), this is realized by 
using the same linkability mechanism as above.
\Ie to forge a witness signature, Alice must produce a \(\wid'\) such that
\begin{multline*}
  \pi_{\wid'}\gets \SPK\left\{ (\sk_W') : \right. \\
    \begin{aligned}
      \wid' &= \ACprf[_{\sk_W'}][\pid] \quad \land \\
      \sigma_W'' &= \left. \ACblind[\ACsign[_{\ssk}][\sk_W']] \right\}
    \end{aligned} \\
      (\pid, \wid', t_s, t_s', l),
\end{multline*}
is accepted by the verifier while not knowing a signature on the secret key 
\(\sk_W'\).
Thus, she must break the \(\AC\) scheme to succeed or get a valid signature on \(\sk_W'\) from the \ac{CA}.
As above, the latter could be done if she can find people willing to collude, 
people unwilling to collude but she can illicitly use their keys anyway or 
simply request new keys from the \ac{CA}.

We assumed that Alice cannot collude with more than a threshold \(\theta\) of 
witnesses.
Thus, setting the threshold of at least \(\theta+1\) will prevent Alice from 
cheating.
(Note that Alice can produce \emph{one} witness signature for herself, since 
she has access to her own \(\sk_P\), but no more.
Thus, \(\theta > 1\) is a requirement in the trust-free case.)

In the trusted witness case, we assume that we can trust (dedicated) witnesses.
In this case, Alice must be able to forge statements, \eg digital signatures, 
from at least one witness that the verifier trusts.

\paragraph{Designated protest}%
\label{analysis-designated}

\Cref{DesignatedEvent} is to prevent Alice (or someone else) from reusing the same proof (or proof share) for another event.
This possibility is prevented through the use of \(\cid\) in the proof shares.
To reuse the proof share for another protest, with a different manifesto, one 
must find a second pre-image \(\mfst'\) such that \(\cid = \Hash[\mfst] = 
  \Hash[\mfst']\).

There exists another case of collision that we must prevent.
Consider the situation in which Alice computes \(\pid = \ACprf[_\sk][\cid]\) for some cause identifier \(\cid\) and some witnesses computes \(\wid_1, \dotsc, \wid_n\), with \(\wid_i = \ACprf[_{\sk_i}][\pid]\).
Now, if Alice constructs a manifesto \(m\) such that \(\Hash[m] = \pid\), then \(\wid_1, \dotsc, \wid_n\) would be valid participant identifiers for the protest with manifesto \(m\).
This means that we must separate the mechanisms used for participating and witnessing.
The protocol achieves this by the fact that \[
  \prf_{\wid} = \SPK[\sk_i][\dotsc][\pid, \wid, t_s, t_s', l]
\]
does not include \(\cid\) whereas \[
  \prf_{\pid} = \SPK[\sk_i][\dotsc][\cid, \pid, \wid, t_s, t_s', l]
\]
does.
Thus, the verification process differs for the two types of proofs.
(An alternative approach would have been to compute them differently, for 
instance \(\pid = \AC[PRF]_{\sk}(x)\) but \(\wid = \AC*[PRF]_{\sk}(x)\), where 
\(\AC \neq \AC*\) are separate anonymous-credential systems.)