### 2019-01-10 paper/reviews/popets-2019.3-rebuttal.md
update of the rebuttal

Reviewer 1

1) a) Due to the use of anonymous credentials, the central authority cannot tell participants from non-participants without extreme measures equivalent to arresting people at the protest. Arguably, the chilling effects with this system are no more than that of "standard" protests. To mitigate the risk of having a central authority responsible the IDs, we could also replace it by a distributed version composed of several independent entities that simulate the work of the central authority is a distributed and private manner.

b) Yes, but the paper also notes that those are detectable and should be ignored due to their obvious bias. In particular, those who verify results must apply the standard source criticism. (e.g., for detecting bias in scientific research).

2) Petitions have been viable for a long time but people still arrange protests anyway, in part because going physically to a protest in a sense shows much more involvment. Our scheme allows for both anonymous verifiable petitions (not considering location) and protests (time and location matter).

3-4) We acknowledge that the design of the architecture is complex but we need all parts to reach the desired security and privacy properties. To improve the clarity of the paper, we agree that we need to present the adversaries better. The objective of introducing A2 was to show the limitations of our approach. While Crocus cannot make physical protests safer, we wanted to ensure that it does make it worse.

5) This is the main reason for distance-bounding as it allows the witness to verify that it is within a given distance from the prover.  Why a verifier accepts this claim is up to the trust assumptions for which we provide two cases. One in which the verifier trusts a particular witness and the other in which the verifier believes that a threshold of witnesses has been reached.

6) Trusted by a verifier. Anyone can define whom they trust and how they want to be counted. We _do_ use trusted witnesses for counting.

7) Strength: yes.

Reviewer 2

If there are specific parts that should be more formalized, we will be happy to do this in the revised version of the paper.

With respect to a single participant gathering proof shares for friends, distance-bounding prevents this to be done by relaying messages. In addition, the terrorist-fraud assumption also prevents a user giving his private key (or phone containing it) to someone. This is realistic in scenario in which the private keys are part of the long-term ID credential. In particular as pointed out by reviewer 1 if this ID can be used for other things, you will not allow just anyone to use it.

Reviewer 3

1) Yes, it follows the standard assumption in distance-bounding to prevent terrorist fraud (which is distance-bounding terminology for this particular case).

Functional encryption will not work against an honest verifier in the general case. We are working on a distance-bounding protocol (in preparation for a separate paper) in which the private key can be computed by joining two different outputs from the functional encryption of the private key, which reduced to the terrorist-fraud case again.

2) This extreme case in which no type of wireless communication works is possible but unheard of. However, it could be potentially solve by using NFC capabilities for instance.

3) There is principled way given online mobs, but different counts are possible. As long as these parameters are published, the counts can be verified. Requests for sizeable collusion are made in public, which risk being found by the verifier --- thus invalidating the count in the verifier's eyes.

4) 5G is built for scale but currently not for the D2D scenario without provider but again NFC is an option. For overall performance, we could evaluate AnonPass on phones, example distance-bounding computation, collect density and scalability from the literature (protests and wireless), and measure NFC delays, which we could do if deemed useful by the reviewers.