### 2019-02-19 paper/reflections-adversary-models.tex
Renames adversary-model discussion to reflections

\section{Reflections on possible adversary models}%
\label{reflections-adversary-model}

% We now provide a concretely defined system and adversary model.
% We give three definitions, each defining an adversary with increasingly stronger 
% capabilities (\ie more auxiliary information).

There are three players: the protester (with identity) \(P\), a witness (with 
identity) \(W\) and the time-stamping storage \(S\) (that will contain the set 
of all proof shares, \(\pshs\)).
The protester \(P\) and the witness \(W\) communicate some protocol data,
\(d_{P,W}(\cid, P)\), and records when the communication occurred \(t_{P,W}\).
The protester \(P\) communicates with \(S\), in which \(S\) only learns some 
function of the protocol data exchanged with the witness, \(f(d_{P,W}(\cid, 
  P))\) for some function \(f\), and the time of the communication 
(\(t_{P,S}\)).
This is illustrated in \cref{fig:base-adversary}.

\begin{figure}
  \centering
  \includegraphics{base-adversary.tikz}
  \caption{\label{fig:base-adversary}%
    An overview of the in-system adversary model.
    The protester with real identity \(P\) and witness with real identity \(W\) 
    communicate.
    They exchange protocol data as a function \(d\) of the protest and protester, \(d_{P,W}(\cid, P)\), and record the time it 
    happened, \(t_{P,W}\).
    The protester submits \(f(d_{P,W}(\cid, P))\), for some function \(f\), to 
    the storage \(S\), who records the time that happened, \(t_{P,S}\).
    Both the witness \(W\) and the storage \(S\) are controlled by the adversary 
    \(A\) (from \cref{in-system-adversary}).
  }
\end{figure}

\begin{adversaries}
\item\label{in-system-adversary}
  \emph{In-system adversary}:
  The adversary maliciously controls \(W\).
  The adversary honest-but-curiously controls \(S\).
  The adversary only learns the protocol data --- \ie what is sent over the 
  channel, no auxiliary data\footnote{%
    By auxiliary data we mean any data outside of the protocol, \ie
    meta-data such as who is on the other side of the channel obtained
    as side information, \eg by face recognition or inference from address
    to identiy.
  }

\item\label{deanonymizing-witness-adversary}
  \emph{Deanonymizing-witness adversary}:
  Everything is the same as in \cref{in-system-adversary}, except that now 
  \(W\) learns \(P\)'s identity from an auxiliary channel.
  However, \(S\) still learns only the protocol data, no auxiliary data.

\item\label{deanonymizing-storage-adversary}
  \emph{Deanonymizing-storage adversary}:
  The situation is the same as in \cref{in-system-adversary}, but now the storage 
  \(S\) learns the protester \(P\)'s identity from an auxiliary channel.
  However, \(W\) still learns only the protocol data, no auxiliary data.

\item\label{deanonymizing-adversary}:
  \emph{Deanonymizing adversary}:
  Everything is the same as in \cref{in-system-model}, except that now both 
  \(W\) and \(S\) learns \(P\)'s identity from an auxiliary channel.
\end{adversaries}

The in-system adversary (\cref{in-system-adversary}) represents an adversary 
that has no access to data from outside the system, \eg inferences that can be 
done from the communication layer (\eg IP-addresses mapped to physical 
identities or face recognition), which means that it has only the protocol data 
at its disposal.

We find \cref{in-system-adversary} suitable when the protester and witness both 
move in a crowd and there is no way for the witness to decide exactly with whom 
he or she communicates with.
(The protesters could also simply wear masks.)
We consider this a fair model for our purposes: if the adversary can 
deanonymize the protesters \emph{at the physical location} and thus arrest 
them, the adversary can do that independently of whether the protesters use our 
protocol or not --- thus our protocol cannot make things worse in such a 
situation.

However, in some situations this might not be the case: \Eg if the crowd is not 
dense the witness will likely see the face of the protester.
If the witness is controlled by the adversary, then it is likely that the 
witness can capture a picture of the face, which can be turned into an identity 
through face recognition.
There are various such scenarios leading to the adversary learning the 
protester's identity, we capture this as 
\cref{deanonymizing-witness-adversary}.

As to \(S\), \cref{in-system-adversary} is equivalent to the adversarial 
storage \(S\) receiving only perfectly anonymous connections.
Another way to see it is that the adversary cannot control \(S\) but can learn 
what \(S\) learns as soon as \(S\) learns it; \eg \(S\) is operated outside the 
adversary's jurisdiction and so the adversary cannot control \(S\), but since 
\(S\) provides public access the adversary can (possibly constantly) query it 
and learn things with little delay.
Posed like this, the adversary cannot observe the channel between \(P\) and 
\(S\), \eg when \(P\) uses tools such as Tor~\cite{Tor}.
Note that the adversary is also allowed to submit data to \(S\) for storage, 
just like everyone else.

\Cref{deanonymizing-storage-adversary} is a much stronger adversary, 
essentially a global, passive adversary.
Even if the adversary cannot observe the channel between \(P\) and \(S\) (as is 
the case with Tor), the adversary can still learn the identity of \(P\) through 
\eg time-correlation attacks (as is the case with Tor).
A nation state can potentially monitor the traffic of all \acp{ISP} and thus 
perform time-correlation attacks on messages on the network and events observed 
from \(S\).
We capture this in \cref{deanonymizing-storage-adversary}, where \(A\) learns 
the identity of \(P\) in the communication with \(S\).

The strongest adversary, \cref{deanonymizing-adversary}, is a combination of 
\cref{deanonymizing-storage-adversary,deanonymizing-wit-adversary}.

\subsection{Security analysis}

We are secure against \cref{in-system-adversary}.

\Cref{deanonymizing-witness-adversary} can link a protester's identity to the 
\(\cid\).
Specifically, the protester proves very strongly that it knows a secret key 
which is used to compute a \(\pid\) based on \(\cid\).

However, we can prevent this attack by hiding the \(\cid\) and blinding the 
\(\pid\) in the \ac{PPK} run with the witness.
We must blind the \(\pid\), otherwise the adversary learns the identity~\(P\) 
corresponding to \(\pid\) from \(W\) and later what \(\cid\) corresponds to 
\(\pid\) from \(S\) --- thus linking \(\cid\) to \(P\).
Hiding \(\cid\) implies that there must be more than one possible \(\cid\), \ie 
that there are both a protest and a counter-protest at the same location.

Hiding the \(\cid\) would also solve the potential problem of biased 
witnessing, \ie that witnesses refuse to act as witnesses for certain 
\(\cid\)'s.

\Cref{deanonymizing-storage-adversary} can link \(\cid\) to \(P\).
Since \(S\) learns the identity \(P\) from the auxiliary channel and learns 
\(\cid\) from the submitted data.
Note that this step cannot use the hiding and blinding above, since that would 
not be verifiable for any third party.

However, we can prevent this attack by submitting proofs in bulk.
If protesters exchange proofs with each other, then they can submit each 
others' proofs --- but this only works if each protester submits proofs for 
more than one \(\cid\)!
This also means that the \ac{NIZK} proofs must be computed during the protest, 
to be exchanged, and not after the protest.

\Cref{deanonymizing-adversary} can always link \(\cid\) to \(P\).
The hiding-and-blinding that protects against \(W\) is not compatible with the 
exchange of unblinded proofs that protects against \(S\).

\subsection{Discussion}

The likelihood of there being both a protest and a counter protest is small, it 
is simply not practiced like that today.
Even if that would be the practice, consider the scenario where skinheads try 
to blend into another crowd.
Thus hiding the \(\cid\) would give minor protection in practice.