### 2019-02-21 paper/privacy-analysis.tex

\subsection{Privacy}

The issue at core from a privacy perspective is linkability.
We must ensure that no part of any proof is linkable to an individual.

We start with \cref{PseudonymUnlink}.
Given \(\pid\), the adversary cannot distinguish whether \(\pid = 
  \ACprf[_{\sk_A}][\cid]\) or \(\pid = \ACprf[_{\sk_B}][\cid]\) due to the 
properties of \(\AC\) (see~\cite{HowToWinTheCloneWars}).

\Cref{ProtestUnlink} means that Alice's proofs must be unlinkable across 
protests.
This also follows from the properties of 
\(\ACprf\)~\cite{HowToWinTheCloneWars}: \(\pid =  \ACprf[_{\sk}][\cid]\) and 
\(\pid' = \ACprf[_{\sk}][\cid']\) (where \(\cid \neq \cid'\)) are unlinkable 
from the perspective of the adversary.
The argument is the same for \cref{WitnessUnlink}.

However, there are more data than \(\pid, \wid\) used in the protocol.
The protocol uses \(\cid, \pid, \wid, t_s, t_s', l, \corr_\pid, \corr_\wid\).
We assume that the \(\cid\) will be used by many individuals, as it is the 
cause identifier.
And thus it cannot be used to uniquely identity any individual.

We assume that the location \(l\) is coarse enough so that many non-overlapping 
\((\pid, \wid)\)-pairs use the same location.
After all, it is the location of the protest, not the location within the 
protest that is interesting.
Thus \(l\) is not uniquely identifying any individual protester or witness.
Likewise, thanks to the constraints in the time-stamp granularity 
(\cref{timestamp}), \(t\) is not uniquely identifying either.

%Overall, informally, the digital trace needed for verifiability in \CROCUS could
%conceivably be used to strengthen a suspicion of participation.
%Given \cref{deanonymizing-storage-adversary} (Grace's) inferences from 
%auxiliary information that points
%to Alice having participated in a protest, \eg pictures at the
%event, with face recognition and subsequent identification, IP address
%of her home router, \etc, then any remaining uncertainty from false
%positives could be reduced by learning from data on the ledger and
%Alice's smartphone. %This potential difference can thus be considered
%                    %the price for verifiability.
%For confirmation of that suspicion, Grace would have to get access to
%and use Alice's private key.