### 2019-05-10 cysep-eurosp/abstract.tex
Adds what's needed for CySeP/Euro-SP abstract

describe the problem being addressed
describe in detail the approach to solve the problem
preliminary results is encouraged but not required
describe the envisioned contribution,
motivate how it relates to the event theme and
why it could spark discussions.

It should also provide information about the presenters and their organizations 
and any special needs for the setup/presentation.