### 2019-05-10 cysep-eurosp/notes.tex
Adds article stuff for cysep-eurosp abstract

\documentclass{article}

\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}

\input{preamble.tex}

\usepackage[noamsthm,notheorems]{beamerarticle}
\setjobnamebeamerversion{slides}

\declaretheorem[numbered=unless unique,style=theorem]{theorem}
\declaretheorem[numbered=unless unique,style=definition]{definition}
\declaretheorem[numbered=unless unique,style=definition]{assumption}
\declaretheorem[numbered=unless unique,style=definition]{protocol}
\declaretheorem[numbered=unless unique,style=example]{example}
%\declaretheorem[style=definition,numbered=unless unique,
%  name=Example,refname={example,examples}]{example}
\declaretheorem[numbered=unless unique,style=remark]{remark}
\declaretheorem[numbered=unless unique,style=remark]{idea}
\declaretheorem[numbered=unless unique,style=exercise]{exercise}
\declaretheorem[numbered=unless unique,style=exercise]{question}
\declaretheorem[numbered=unless unique,style=solution]{solution}

\begin{document}
\title{%
  Something
}
\author{Daniel Bosk and Sonja Buchegger\\
KTH Royal Institute of Technology}
\institute{%
  KTH EECS
}

\mode<article>{\maketitle}
\mode<presentation>{%
  \begin{frame}
    \maketitle
  \end{frame}
}

\mode*

\mode<all>{\input{abstract.tex}}

%%% REFERENCES %%%

\begin{frame}[allowframebreaks]
  \printbibliography
\end{frame}
\end{document}