### 2019-05-10 cysep-eurosp/abstract.tex
adds some sections in abstract

% describe the problem being addressed
% describe in detail the approach to solve the problem
% preliminary results is encouraged but not required
% describe the envisioned contribution,
% motivate how it relates to the event theme and
% why it could spark discussions.

% It should also provide information about the presenters and their organizations 
% and any special needs for the setup/presentation.

\textbf{Problem:}
Crowd counts vary according to who is doing the counting and what
methods and criteria they use. Current methods have wide margins for
error, are difficult to verify, and can be privacy-invasive. We
recognize the potential adversarial setting for crowd counts and thus
aim for both transparency (in the form of verifiability) and
privacy. Different counts: Any solution

\textbf{Approach:}
 In this paper we propose CROCUS, a decentralized system based on
 smartphones that combines anonymous credentials, witnesses of
 proximity, and storage of participation proofs on a public ledger
 (e.g., blockchain) to achieve properties similar to those needed for
 electronic voting, where, put simply, it is important to count
 everyone, but only once, not leak private information, and be able to
 prove that the count was done correctly.

\textbf{More detailed approach:} 
By running a distance bounding scheme with anonymous credentials,
physically present persons (such as protest participants) and create proof
shares. This is done using smartphones with (assumed)
distance-bounding chips, without Internet connection but using any
available local communications. Before and after the event (protest), participants (and
witnesses), interact with a public ledger such as a blockchain (which does require
Internet access, and anonymization by e.g. Tor). Before the event to get
a timestamp and after the event to upload proof shares, again with a
time stamp. The proof shares form verifiable yet privacy-preserving
participation proofs according to counting eligibility criteria. These criteria, including
time interval, location range, number or kind (e.g., anyone, trusted
journalists) of unique proof shares
needed for a proof, enable verification by recount. Counting is done
by downloading the proof shares from the ledger, checking the validity
and eligibility, and summing up. Any published counts need to also
publish the criteria used to be verified, but anyone could decide on
their own criteria and count accordingly.

\textbf{Preliminary results:} We find that, with some
 assumptions about the availability of future technology such as
 distance-bounding chips on smartphones, CROCUS can provide the
 desired properties at acceptable performance.

\textbf{Relation to theme:} This work is concerned with both security
and privacy: accountability and verifiability (individual, universal, eligibility) and
privacy (unlinkability to individual, unlinkability between events,
unlinkability between witnessing instances).

\textbf{Why it could spark discussions:}
practical/ideological problems: sybil-proof identities needed for anonymous
credentials

\textbf{Information about the presenters:} Daniel Bosk is a PhD
student and Sonja Buchegger a Computer Science professor, both at the
division of Theoretical Computer Science at KTH Royal Institute of
Technology, working on privacy-enhancing technologies.