### 2020-03-13 paper/ZKPK.tex
Comments out SPK (not used in this paper)

\subsection{Zero-knowledge proofs of knowledge}%
\label{ZKPK}

We will use the notation introduced by \citet{PKnotation}:
\begin{equation}\label{PKexample}
  \PK[\alpha, \beta, \gamma][y = g^\alpha h^\beta \land y' = \hat g^\gamma],
\end{equation}
is \iac{PK} protocol where the prover proves knowledge of \(\alpha, \beta, 
\gamma\) ensuring that \(y, y'\) are of the form \(y = g^\alpha h^\beta\) and 
\(y' = \hat g^\gamma\), respectively.
Greek letters are known only to the prover and used for the information for which the prover wishes to prove knowledge, while all other letters are known by the verifier.

%When a proof of knowledge is turned into a signature using the Fiat-Shamir heuristic~\cite{FiatShamirHeuristic}, we will denote it as
%\begin{equation*}\label{SPKexample}
%  \sigma\gets \SPK[\alpha, \beta, \gamma]%
%  [y = g^\alpha h^\beta \land y' = \hat g^\gamma]%
%  [m],
%\end{equation*}
%(Signature based on Proofs of Knowledge) which yields a signature \(\sigma\) on 
%\(m\), ensuring that the issuer knows \(\alpha, \beta, \gamma\) such that \(y = 
%g^\alpha h^\beta\) and \(y' = \hat{g}^\gamma\).