### 2020-03-13 paper/conclusions.tex
Improves conclusion scenario (fixes #46)

\section{Conclusion}%
\label{Conclusion}

In this paper, we have adapted the Schnorr protocol as a
distance-bounding protocol.  It allows to construct distance-bounding
general attribute-based authentication, more general attributes than
provided by previous \ac{DB} ones.  With our contribution, Alice can
now prove more complex statements to Bob; \eg that she is older than 18 
(without revealing her actual age), has a cinema subscription which is not 
\enquote{double spent} by someone else at the same time and, at the same time, 
Bob can be sure that Alice is not just relaying messages from her older sister 
Carol or \ac{DBMF} relaying the subscription from some unsuspecting stranger.