### 2020-07-03 paper/conclusions.tex
Edits conclusions

\section{Conclusion}%
\label{Conclusion}

In this paper, we have adapted the Schnorr protocol as \iac{DB} protocol.
It allows us to construct \ac{DB}, general attribute-based authentication; more 
general attributes than provided by previous \ac{DB} ones.
With our contribution, Alice can now prove more complex statements to Bob; \eg 
that she is older than 18 (without revealing her actual age), has a cinema 
subscription which is not \enquote{double spent} by someone else at the same 
time~\cite[\eg][]{AnonPass} and, at the same time, Bob can be sure that Alice 
is not just relaying messages from her older sister Carol or \ac{DBMF} relaying 
the subscription from some unsuspecting stranger also waiting in the queue to 
enter.