### 2021-06-16 paper/related-work.tex
Updates relatd-work.tex with broken out parts

\section{Related work}%
\label{related-work}

Our scheme relates to several other works: crowd counting methods, location 
proof systems, and Sybil-free pseudonym systems.

\input{current-crowd-counting}
\input{location-proof-systems}
\input{sybil-free-systems}