### 2021-06-16 paper/practical-considerations.tex

\section{Practical Considerations}%
\label{Practical}

We now briefly list some issues for deploying
CROCUS, regarding implementability and realistic adversaries\footnote{see the long version of the paper for a more thorough
discussion of these topics. }, and then analyze its performance.

\subsection{Implementability}

Some of the assumptions that are required for implementing our proposition are 
not yet realized, however, we believe that they soon will be:
ubiquitous usage of smartphones, with an identity credential
signed by some \ac{CA}\footnote{based on \eg governmental ID on a chip as in Estonia, national
  but private as BankID in Sweden, eIDAS in the EU, or self-sovereign
  identities} ,  distance-bounding capabilities\footnote{based on \eg
  3DB-access or key-fob technology integration with phones}, and
infrastructure-independent wireless communications\footnote{based on
  \eg NFC (available but inconvenient), bluetooth or WIFI (Briar and
  FireChat have been used in protests), 5G D2D (scalable but needs
  authorization step)}.

\subsection{Realistic adversary model}
Our adversary model considers only protocol data, no auxiliary data.
Against this adversary our scheme is secure.
The question is how this adversary model maps to real
adversaries. Side-channels such as taking photos, a global passive adversary
controlling all national ISPs doing time-correlation attacks (though
may not be able to observe all ledger input or \eg Tor exit nodes), or
protocol violations in protests organized by the \ac{CA} are
not prevented.

\input{performance-subsection}