### 2021-06-17 paper/related-work.tex

\section{Related work}%
\label{related-work}

Our scheme relates to several other works: crowd counting methods, location 
proof systems, and Sybil-free pseudonym systems.

\input{current-crowd-counting-short}
\input{location-proof-systems-short}
\input{sybil-free-systems}