### 2021-06-17 paper/location-proof-systems-short.tex

\paragraph{Location-proof systems}%
\label{location-proof-systems}

Our work is very related to \acp{LPS}~\cite{luo2010veriplace,zhu2011applaus}, a digital certificate attesting that someone was at 
a particular location at a specific moment in time, except we need to
tie the cause of the protest to the location proof and cannot trust
the verifier. Most of the existing  \acp{LPS} approaches require the prover and the 
witnesses to disclose their identities, thus raising many privacy
issues.
PROPS~\cite{PROPS} is  an exception, but like PASPORT~\cite{PASPORT}
and \Textcite{ProofOfWitnessPresence}, does not treat the adversarial
setting needed for \CROCUS.