### 2021-06-17 paper/distance-bounding.tex
Cuts distance bounding further

\subsection{Distance-bounding protocols}%
\label{distance-bounding}

\Ac{DB} protocols were first suggested by \citet{DistanceBounding} to prevent relay attacks in contactless communications in which the adversary forwards a communication between a prover and a possibly far-away verifier to authenticate. 
These attacks cannot be prevented by cryptographic means as they are independent of the semantics of the messages exchanged.
As a consequence, mechanisms ensuring the physical proximity between a verifier and a prover should be used instead.
\Ac{DB} protocols precisely enable the verifier to estimate an upper bound on 
their distance to the prover during authentication.
%by measuring the time-of-flight of short challenge-response messages (or 
%rounds) exchanged during time-critical phases. Time critical phases are 
%complemented by slow phases during which the time is not taken into account. At 
%the end of a \Ac{DB} protocol, the verifier should be able to
%determine if the prover is legitimate \emph{and} in their vicinity.
%In this sense, \Ac{DB} protocols combine the classical properties of authentication protocols with the possibility of verifying the physical proximity.
%
Our setting requires a public-key \ac{DB} protocol with a \emph{malicious 
verifier} who will potentially try to \emph{impersonate the prover}.
The verifier might also try to track the provers and map their identities to 
their actions, thus we also require strong privacy.
There exists one such protocol in the literature, namely one by 
\textcite{DB-Schnorr}, we refer to that paper for a detailed discussion.