### 2021-06-17 paper/current-crowd-counting-short.tex
shortens current crowd counting

\paragraph{Current crowd-counting methods}

The most common method for counting crowds at protests is \emph{Jacobs's method}~\cite{2016DemonstrationsInSeoul,BBCHowToCountProtestNumbers}, estimating crowd density in different regions seen from aerial pictures and summing up.  Using pictures makes it difficult to get cumulative counts, verify that the pictures have not been manipulated, and to have both privacy and individual verifiability: either one is included in the picture (privacy problem) or not (verifiability problem). Similar limitations exist for estimating the number of persons in a picture or video (\eg CrowdSize~\cite{CrowdSize}).
Edge cases are difficult: determining whether people near the event's perimeter are participants or 
simply bystanders~\cite{HowToEstimateCrowdSize}, and as noted with demonstrations in Seoul:
\blockcquote{2016DemonstrationsInSeoul}{%
  \textins*{t}he demonstrators not only gather in open space, but also small alleys and between buildings%
}. Counting MAC addresses, as done by a company there, suffers from MAC randomization, though some tracking of smartphones could still be possible with a different method~\cite{WhyMACRandomizationIsNotEnough} or using \emph{IMSI catchers}; none of which is verifiable.
Trusted infrastructure was deployed by a collection of media outlets to count protesters passing the line defined by a trusted sensor on marches~\cite{LeMondeProtestingSolution}. Again, due to the lack of
strong verifiability guarantees it was complemented by micro-counts made by humans. %to estimate their margin of error.

% \Textcite{CrowdCount} is a web service that lets Alice create an event such 
% that anyone can submit their location to register that they are in Alice's 
% event.
% Another related approach based on devices is UrbanCount~\cite{UrbanCount}, which relies on epidemic spreading of crowd-size estimates by device-to-device communication to count crowds in dense urban environments with high node-mobility and churn.
% However, there is no consideration of a potentially adversarial setting and thus no verifiability or checks on eligibility.  DiVote~\cite{DiVote}, a prior work by the same authors for polling in dense areas, avoids double counting, but again only works with honest participants and thus does not suit an adversarial setting.